### 2020-11-02 MSE327/EDA_outcome.Rmd
EDA script for PM2.5 histogram

---
title: "MSE 327 Project"
author: "Davide Giovanardi"
date: "11/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
setwd("/Users/davidegiovanardi/Desktop/MSE327/PM2.5-Nonattainment/MSE327/")
load("pmanalysis.RData")
```

# EDA
## Outcomes
### Ambient PM2.5 2010-2012

```{r EDA}
# Check that it matches Table 1
cat(mean(na.omit(dat[dat$a == 0, "pmfu"])))
cat(mean(na.omit(dat[dat$a == 1, "pmfu"])))

clean_df <- function(dat, outcome) {
  df <- na.omit(dat[,c("a",outcome)])
  return(data.frame(df))
}

df <- clean_df(dat, "pmfu")

df$Area <- c("attainment", "non-attainment")[(df$a)+1]

ggplot(df, aes(pmfu, color = Area, fill = Area)) +
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5, bins = 25) +
  geom_density(alpha = 0.25) +
  theme_minimal() + 
  ggtitle("Outcome: Ambient PM2.5 2010-2012")

```