### 2020-11-08 Analysis Files/Paper_Analysis_327.Rmd
adding our paper analyisis rmd file

---
title: "Summary of PM2.5 Nonattainment  Analysis"
author: "Cory Zigler"
date: "December 20, 2016"
output: pdf_document
---

```{r}
rm(list=ls())
set.seed(51)
# library(HEIfunctions)
library(maps)
library(PSAgraphics)
library(data.table)

source('mybalcheck.R')
source('ReadSimsFunction.R')
getrate=function(vec, denom){
  return(vec*1000/denom)
}

#########   Read in the data and define baseline and follow up measures    ##########
#### Read in the Data ######
# --- 'pmanalysis.RData' is a file created from the data construction files
# ---  Use this file if you want to see the whole analysis from start to finish
# ---  But note that the data sets and results produced with this file will be different
# --- from those reported in the paper (sample size, propensity score estimates, effect estimates, etc.).
# ---  This is because all of the Medicare data are simulated
load(file="pmanalysis.RData")

#### Read in the Data ######
# --- 'pm_withps_nomed.Rda' is a file that is provided that is exactly the same as that
# --- used for the paper, except with all Medicare variable set to zero.  Use this file if
# --- you wish to reproduce the basic summaries (not involving Medicare), sample sizes, and Table 1.  This 
# --- file has the actual propensity sore estimates used in the paper.
# --- Note that this use of this file wil preclude the entirety of this script from running, because 
# --- the code after line 166 relies on the Medicare data
# load(file="pm_withps_nomed.Rda")

datorig = dat

dat$pmbase = dat$pmbase2002_2004
dat$baseorig = dat$baseorig2002_2004

#########   Say which results you will use for the final analysis   #####
pollutiontransform = "log"

whichresults = "fake_med/"
whichresults_noprune = "fake_med_noprune/"
whichresults_pollution = "fake_poll_med/"
whichresults_pollution_noprune = "fake_poll_med_noprune/"

# Specify C for summarizing average associative and dissociative effects
Cede= 1.0
Ceae = -1.0

nburn = 10
thin = 1

outcomes = c("total_death_FFS_MA.2012", "COPD.2012", "CV_stroke.2012", "HF.2012", "HRD.2012", "IHD.2012",
                 "PVD.2012", "RTI.2012")
outcomes.2004 = c("total_death_FFS_MA.2004", "COPD.2004", "CV_stroke.2004", "HF.2004", "HRD.2004", "IHD.2004",
                 "PVD.2004", "RTI.2004")
dat$mortrate = 1000*dat$total_death_FFS_MA.2004/dat$Tot_den_for_death_MA_FFS.2004
dat$CVDrate = 1000*dat$ALL_CVD.2004/dat$Person_year_FFS.2004
dat$resprate = 1000*(dat$COPD.2004 + dat$RTI.2004)/dat$Person_year_FFS.2004
dat$COPDrate = 1000*dat$COPD.2004/dat$Person_year_FFS.2004
dat$CVstrokerate = 1000*dat$CV_stroke.2004/dat$Person_year_FFS.2004
dat$HFrate = 1000*dat$HF.2004/dat$Person_year_FFS.2004
dat$HRDrate = 1000*dat$HRD.2004/dat$Person_year_FFS.2004
dat$IHDrate = 1000*dat$IHD.2004/dat$Person_year_FFS.2004
dat$PVDrate = 1000*dat$PVD.2004/dat$Person_year_FFS.2004
dat$RTIrate = 1000*dat$RTI.2004/dat$Person_year_FFS.2004
outcomes.2004.rate = c("mortrate", "CVDrate", "resprate", "COPDrate", "CVstrokerate", "HFrate", "HRDrate", "IHDrate", "PVDrate", "RTIrate")
```

### Basic Data Description (Section 2.2)
```{r,  echo=FALSE}
print(paste("Total number of locations =", dim(dat)[[1]]))
print(paste("Total number of fee-for-service Medicare beneficiaries =", sum(dat$Total_den_FFS.2012)))
print(paste("Total number of managed care Medicare beneficiaries =", sum(dat$Tot_den_for_death_MA_FFS.2012) - sum(dat$Total_den_FFS.2012)))
```
### Nonattainment Description (Section 2.2)
```{r,  echo=FALSE}
print(paste("Number of nonattainment areas =", sum(dat$a==1)))
print(paste("Number of nattainment areas =", sum(dat$a==0)))

whichyears = 1999:2012  ## only 3 monitors in 1997, only 4 in 1998
whichvar = "Arithmetic.Mean" ##"X95th.Percentile" ##Arithmetic.Mean"
attaindat=read.csv('Data Construction/Other Data Files/EPA Nonattainment Table.csv')
attaindat$FIPS=paste(formatC(attaindat$fips_state, width=2, flag="0"), formatC(attaindat$fips_cnty, width=3, flag="0"), sep="")
attaindat = subset(attaindat, pollutant=='PM-2.5 1997')
attaindat$a_2012 = ifelse(attaindat$pw_2012 %in% c("P", "W"), 1, 0)
checka_2012 = attaindat[, c("FIPS", "a_2012")]
check_attain_fu = merge(dat, checka_2012, by = "FIPS", all.x = TRUE)
check_attain_fu$a_2012[is.na(check_attain_fu$a_2012)] = 0
with(check_attain_fu, table(a, a_2012))
print(paste("Percentage of 2005 nonattainment areas that remained in nonattainment in 2012 =", with(check_attain_fu, round(sum(a_2012==1)/(sum(a==1)),2))))

```

### Covariate Description in Attainment and Nonattainment Areas (Table 1)
```{r, echo=FALSE}
covariates = c("pmbase", "a_ozone",                                                                    ## AQS variables
               "TotPop", "logpop", "CompletelyRural", "PctUrban",  "PctBlack", "PctHisp", "PctHighSchool", "MedianHHInc",  ## Census Variables
               "PctPoor", "PctFemale", "PctOccupied", "PctMovedIn5", "MedianHValue",  
               "smokerate2000",                                                                        ## Smoking rate
               "avgdewpt", "avgtemp", "avgrelhum",                                                     ## Weather variables       
               "Tot_den_for_death_MA_FFS.2004", "mean_age.2004", "Female_rate.2004",                   ## Medicare variables
               "White_rate.2004", "Black_rate.2004", outcomes.2004.rate)    
othervars = c("Monitor", "Longitude", "Latitude", "a")

###### Create "Table 1" datatab ######
# --- "*" for variables included in the PS model$^*$
# --- "+" for variables included for adjustment in outcome model $^+$
# --- updated 20161209
tabvars = covariates[!(covariates %in% c("TotPop"))]
varnames = c("Ambient PM2.5 2002-2004 ($\\mu g/m^3$)$^*$$^+$", "\\% Ozone nonattainment 2005$^*$", ### Monitor-Level
           "log(population)$^*$$^+$", "Completely Rural Area$^*$", "\\% Urban$^*$$^+$", "\\% Black$^+$", "\\% Hispanic$^+$", "\\% HS Grad.$^*$$^+$", "Median HH Inc. (\\$)$^*$",
           "\\% Poor$^*$$^+$", "\\% Female", "\\% Occupied Housing$^*$$^+$", "5-Year Migration Rate$^*$", "Median House Value (\\$)$^*$", 
           "Smoking Rate$^*$", "Avg. Dew Point ($^\\circ$F)$^*$", "Avg. Temperature ($^\\circ$F)$^*$", "Avg. Rel. Humidity (\\%)$^*$$^+$",
           "Total Medicare Benef. 2004$^*$$^+$", "Avg. Medicare Age 2004 (years)$^*$$^+$", "\\% Female Medicare Benef. 2004$^*$$^+$", "\\% White Medicare Benef. 2004$^*$$^+$",
           "\\% Black Medicare Benef. 2004$^*$$^+$",
           "Mortality$^*$", "All CVD$^*$", "Respiratory$^*$", "COPD", "CV Stroke", "Heart Failure", "HRD", "Ischemic Heart Disease", "Peripheral Vascular Disease", "Respiratory Tract Infection")  ## $^*$ for in PS, $^+$ for in outcome too
varnames_plot = c("Ambient PM2.5 2002-2004", "\\% Ozone nonattainment 2005", ### Monitor-Level
           "log(population)", "Completely Rural Area", "% Urban", "% Black", "% Hispanic", "% HS Grad.", "Median HH Inc.",
           "% Poor", "% Female", "% Occupied Housing", "5-Year Migration Rate", "Median House Value", 
           "Smoking Rate", "Avg. Dew Point", "Avg. Temperature", "Avg. Rel. Humidity",
           "Total Medicare Benef.", "Avg. Medicare Age 2004", "% Female Medicare Benef. 2004", "% White Medicare Benef. 2004",
           "% Black Medicare Benef. 2004",
           "Mortality", "All CVD", "Respiratory", "COPD", "CV Stroke", "Heart Failure", "HRD", "Ischemic Heart Disease", "Peripheral Vascular Disease", "Respiratory Tract Infection")

healthnames = c("mort",  "copd", "cvstroke", "hf", "hrd", "ihd", "pvd", "rti")    
names(healthnames) = c("total_death_FFS_MA.2012",
                       "COPD.2012", "CV_stroke.2012", "HF.2012", "HRD.2012", "IHD.2012",
                       "PVD.2012", "RTI.2012") 
denominators = c("Tot_den_for_death_MA_FFS.2012", rep("Person_year_FFS.2012", length(healthnames)-1))
for (h in 1:length(healthnames))
  dat[, paste(healthnames[h], "rate", sep = "_"):= 1000*dat[, names(healthnames)[h], with = FALSE]/dat[, denominators[h], with = FALSE]]

outvars = c("pmfu", paste(healthnames, "rate", sep = "_"))
outnames = c("Ambient PM2.5 2010-2012 ($\\mu g/m^3$)", "Mortality 2012", "COPD 2012", "CV Stroke 2012", "Heart Failure 2012", "HRD 2012", 
             "Ischemic Heart Disease 2012", "Peripheral Vascular Disease 2012", "Respiratory Tract Infection 2012")

allvars = c(tabvars, outvars)
allnames = c(varnames, outnames)
vartab = matrix(NA, nrow=length(allvars), ncol=4)
dimnames(vartab)[[1]] = allnames
dimnames(vartab)[[2]] = rep(c("Mean", "SD"), 2)

for (i in 1:length(allvars) ){
  vartab[i,c(1,3)] = with(dat, tapply(as.matrix(dat[,allvars[i], with = FALSE]), a, mean, na.rm=TRUE))
  vartab[i,c(2,4)] = with(dat, tapply(as.matrix(dat[,allvars[i], with = FALSE]), a, sd, na.rm=TRUE))
  #  vartab[i,3] = t.test(dat[, vars[i]]~dat[, "a"])$p.value
}

## convert decimals to percentages
whichpercentvars = which(substr(dimnames(vartab)[[1]], 1, 2) == "\\%")
# whichpercentvars = c(2,4,5,6,7,9,10,11,20,21,22)
vartab[whichpercentvars,] = 100*vartab[whichpercentvars,]
print(round(vartab,2))
# write.table(round(vartab,2), file = "vartab.txt", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=FALSE, quote=FALSE)
```

# Propensity Score Analysis
```{r, echo=FALSE}
###############################################################
#####     Estimate the Propensity Score      ##################
###############################################################
dat$CompletelyRural = as.numeric(dat$CompletelyRural)
# -- Full model using only summary health outcomes
fullmodel = glm(a~.  + avgrelhum*avgtemp + pmbase*avgtemp, family=binomial(link="logit"), data = dat[, names(dat) %in% c(covariates, "a") & !(names(dat) %in% c("Monitor", "TotPop", "COPDrate", "CVstrokerate", "HFrate", "HRDrate", "IHDrate", "PVDrate", "RTIrate")), with = FALSE]) #"mortrate", "resprate", "CVDrate"

# - Take out Census covariates redundant with Medicare
allvars = rownames(summary(fullmodel)$coef)
inclvars = allvars[!(allvars) %in% c("PctBlack", "PctHisp", "PctFemale")]
inclvars = inclvars[!inclvars %in% c("(Intercept)")]
inclformula = as.formula(paste("a ~", paste(inclvars, collapse = "+")))
psmod = glm(inclformula, family=binomial(link="logit"), data = dat) 
print("The propensity score model summary:")
summary(psmod)

dat$ps = round(psmod$fitted.values, 3)
mina1 = min(dat$ps[dat$a==1])
maxa0 = max(dat$ps[dat$a==0])
dat$outofrange = 0

print(paste("Minimum estimated propensity score =", min(dat$ps)))
print(paste("Maximum estimated propensity score =", max(dat$ps)))

print(paste("Minimum propensity score among nonattainment areas =", mina1))
print(paste("Maximum propensity score among attainment areas =", maxa0))

###### Prune based strictly on range of observed PS  ######
dat$outofrange[dat$ps>maxa0 & dat$a==1]=1
dat$outofrange[dat$ps<mina1 & dat$a==0]=1

##### Define PS categories ##################
cuts = quantile(dat$ps[dat$outofrange==0], seq(.25, .75, .25)) # quartiles
dat$pscat[dat$outofrange==0] = as.numeric(cut(dat$ps[dat$outofrange==0], breaks=c(0,cuts,1), include.lowest=TRUE))
dat$pscat = as.factor(dat$pscat)

print(paste("Number of nonattainment areas pruned =", sum(dat$outofrange==1 & dat$a==1)))
print(paste("Number of attainment areas pruned =", sum(dat$outofrange==1 & dat$a==0)))

pstab = with(dat, table(a,pscat))
pstab = rbind(pstab, with(dat, table(pscat)))
print(pstab)
write.table(pstab, file = "pstab.txt", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=FALSE, quote=FALSE)

printvars = tabvars
covnames = varnames_plot
length(printvars) == length(covnames)

# pdf("balplot.pdf", height=10, width=10)
par(mar = c(5,10,2,2))
bal = mybalcheck(covariates=as.matrix(dat[, printvars, with = FALSE]), treatment=dat$a, propensity=dat$ps, strata=dat$pscat, absolute.es=FALSE, xlim=c(-1,2), outofrange=dat$outofrange, universal.psd=FALSE, plot.strata=FALSE, covnames = covnames)
# dev.off()

#### which variables have big differences after PS adjustment (residual confounding)
# -- Do this in the pruned sample for the main analysis
stratabal = abs(bal$effect.sizes[, c("a", "b", "c", "d")])

### output a list of variables to include in the models for residual adjustment 
#### include if: 1) average bal > 0.25 or any stratum greater than 0.5
whichresid = matrix(0, nrow = dim(stratabal)[[1]], ncol = dim(stratabal)[[2]])
dimnames(whichresid) = dimnames(stratabal)
residtoobig = ( abs(stratabal) > 0.5)
whichresid[residtoobig] = stratabal[residtoobig]
checkresid = cbind(whichresid, apply(stratabal, 1, mean))
includevar = abs(checkresid[, dim(stratabal)[[2]]+1]) > 0.25 | rowMeans(abs(checkresid[, 1:(dim(stratabal)[[2]])]) ) > 0

cbind(checkresid, includevar)
varnames = dimnames(checkresid)[[1]][includevar]
vars_with_imbalance = printvars[(covnames %in% varnames) & !(covnames %in% c("Mortality", "All CVD", "Respiratory", "COPD", "CV Stroke", "Heart Failure", "HRD", "Ischemic Heart Disease", "Peripheral Vascular Disease", "Respiratory Tract Infection"))]
save(vars_with_imbalance, file = "vars_with_imbalance.RData")
print("Variables with residual imbalance that are included for direct covariate adjustment:")
print(vars_with_imbalance)

## Output data set with propensity scores
save(dat, file = "pm_withps.Rda")

fulldat = dat
```

## Maps of Monitoring Locations (Figure 1)
```{r,  echo=FALSE}

## --- Propensity Score Maps ----  ##
colors = c(rgb(0,0,1,.75), rgb(1,0,0,.5))

## Restrict to Eastern US
whichstates_abb = c("MO", "AR", "IA", "IL", "IN", "MI", "KY", "TN", "MS", "AL", "GA", "SC", "NC", "VA", "WV", "OH", "PA", "DE","DC", "NY", "MA", "CT", "MD", "NH", "VT", "WI", "LA", "MN", "FL", "ME") 
whichstates_fips = state.fips[state.fips$abb %in% whichstates_abb, "fips"]

# -- All locations (Figure 1a)
# pdf("monitormap_all.pdf")
map("state", region = state.fips[state.fips$abb %in% whichstates_abb, "polyname"])
with(dat, points(Longitude, Latitude, pch = 16, cex=1, col = colors[a+1]))
legend("bottomleft", pch=16, cex=1, col=colors, c("Attainment", "Nonattainment"), bty="n")
# dev.off()

# -- Retained locations (Figure 1b)
# pdf("monitormap_pruned.pdf")
map("state", region = state.fips[state.fips$abb %in% whichstates_abb, "polyname"])
with(subset(dat, outofrange==0), points(Longitude, Latitude, pch = 16, cex=1, col = colors[a+1]))
legend("bottomleft", pch=16, cex = 1, col=colors, c("Attainment", "Nonattainment"), bty="n")
# dev.off()

```

## Table Comparing Pruned to Retained (Appendix Table XX)
```{r,  echo=FALSE}
varnames = c("Ambient PM2.5 2002-2004 ($\\mu g/m^3$)$^*$$^+$", "\\% Ozone nonattainment 2005$^*$", ### Monitor-Level
           "log(population)$^*$$^+$", "Completely Rural Area$^*$", "\\% Urban$^*$$^+$", "\\% Black$^+$", "\\% Hispanic$^+$", "\\% HS Grad.$^*$$^+$", "Median HH Inc. (\\$)$^*$",
           "\\% Poor$^*$$^+$", "\\% Female", "\\% Occupied Housing$^*$$^+$", "5-Year Migration Rate$^*$", "Median House Value (\\$)$^*$", 
           "Smoking Rate$^*$", "Avg. Dew Point ($^\\circ$F)$^*$", "Avg. Temperature ($^\\circ$F)$^*$", "Avg. Rel. Humidity (\\%)$^*$$^+$",
           "Total Medicare Benef. 2004$^*$$^+$", "Avg. Medicare Age 2004 (years)$^*$$^+$", "\\% Female Medicare Benef. 2004$^*$$^+$", "\\% White Medicare Benef. 2004$^*$$^+$",
           "\\% Black Medicare Benef. 2004$^*$$^+$",
           "Mortality$^*$", "All CVD$^*$", "Respiratory$^*$", "COPD", "CV Stroke", "Heart Failure", "HRD", "Ischemic Heart Disease", "Peripheral Vascular Disease", "Respiratory Tract Infection")  ## $^*$ for in PS, $^+$ for in outcome too
allvars = c(tabvars, outvars)
allnames = c(varnames, outnames)
vartab = matrix(NA, nrow=length(allvars), ncol=8)
dimnames(vartab)[[1]] = allnames
dimnames(vartab)[[2]] = rep(c("Mean", "SD"), 4)

for (i in 1:length(allvars) ){
  # - table columns are Attainment retained, Attainment pruned, Nonattainment retained, Nonattainment pruned
  vartab[i,c(1,3)] = with(subset(dat, a==0), tapply(as.matrix(dat[a==0,allvars[i], with = FALSE]), outofrange, mean, na.rm=TRUE))
  vartab[i,c(2,4)] = with(subset(dat, a==0), tapply(as.matrix(dat[a==0,allvars[i], with = FALSE]), outofrange, sd, na.rm=TRUE))
  vartab[i,c(5,7)] = with(subset(dat, a==1), tapply(as.matrix(dat[a==1,allvars[i], with = FALSE]), outofrange, mean, na.rm=TRUE))
  vartab[i,c(6,8)] = with(subset(dat, a==1), tapply(as.matrix(dat[a==1,allvars[i], with = FALSE]), outofrange, sd, na.rm=TRUE))
  #  vartab[i,3] = t.test(dat[, vars[i]]~dat[, "a"])$p.value
}

## convert decimals to percentages
whichpercentvars = which(substr(dimnames(vartab)[[1]], 1, 2) == "\\%")
# whichpercentvars = c(2,4,5,6,7,9,10,11,20,21,22)
vartab[whichpercentvars,] = 100*vartab[whichpercentvars,]
print(round(vartab,2))
write.table(round(vartab,2), file = "vartab_pruned_v_retained.txt", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=FALSE, quote=FALSE)

```

# Results (Section 3)
```{r,  echo=FALSE}
print(paste("Total number of locations retained after pruning =", sum(dat$outofrange==0)))
print(paste("Total number of fee-for-service Medicare beneficiaries in pruned sample = ", sum(dat$Total_den_FFS.2012[dat$outofrange==0])))

print(paste("Total number of managed care Medicare beneficiaries in pruned sample = ", sum(dat$Tot_den_for_death_MA_FFS.2012[dat$outofrange==0]) - sum(dat$Total_den_FFS.2012[dat$outofrange==0])))

with(subset(dat, outofrange==0), table(a))
print(paste("The number of nonattainment locations in the pruned sample =", with(subset(dat, outofrange==0), sum(a==1))))
```

### Missing Data Description (Appendix C.3)
```{r,  echo=FALSE}
print(paste("The number of locations with missing baseline pollution =", sum(is.na(dat$baseorig))))
print(paste("The number of nonattainment locations with missing baseline pollution = ", sum(is.na(dat$baseorig) & dat$a==1)))
print(paste("The number of retained locations with missing baseline pollution =", sum(is.na(dat$baseorig) & dat$outofrange==0)))
print(paste("The number of retained nonattainment locations with missing baseline pollution =", sum(is.na(dat$baseorig) & dat$outofrange==0 & dat$a==1)))

print(paste("The number of locations with missing follow up pollution =", sum(is.na(dat$pmfu))))
print(paste("The number of nonattainment locations with missing follow up pollution = ", sum(is.na(dat$pmfu) & dat$a==1)))
print(paste("The number of retained locations with missing follow up pollution =", sum(is.na(dat$pmfu) & dat$outofrange==0)))
print(paste("The number of retained nonattainment locations with missing follow up pollution =", sum(is.na(dat$pmfu) & dat$outofrange==0 & dat$a==1)))
```

## Pollution Analysis Results
```{r, echo=FALSE}
dat = subset(fulldat, outofrange==0)

resultname = paste(whichresults_pollution, "pollutionmodel.RData", sep = "")
load(resultname)

nsamp = dim(mod$samples)[[1]] - 1 ## "-1" used because last entry is NA

y0out = exp(mod$y0[seq(nburn,nsamp,thin),])
y1out = exp(mod$y1[seq(nburn,nsamp,thin),])

ceymat = y1out-y0out
atty = rowMeans(ceymat[, dat$a==1])

print(paste("The posterior mean effect on ambient PM2.5 = ", round(mean(atty), 3)))
print(paste("with 95% posterior interval = (", round(quantile(atty, c(.025)), 3), ",", round(quantile(atty, c(.975)),3), ")"))
print(paste("The posterior probability that ATTy < 0 = ", round(sum(atty<0)/length(atty),3)))
```

## Analysis of Health Outcomes
```{r, echo=FALSE}
healthnames = c("mort", "copd", "cvstroke", "hf", "hrd", "ihd", "pvd", "rti")    
names(healthnames) = c("total_death_FFS_MA.2012",
                       "COPD.2012", "CV_stroke.2012", "HF.2012", "HRD.2012", "IHD.2012",
                       "PVD.2012", "RTI.2012") 
fullhealthnames = c("Mortality", "COPD", "CV Stroke", "Heart Fail.", "Heart Rhythm Disord.", "IHD", "PVD", "Resp. Tract Infect.")
denominators = c("Tot_den_for_death_MA_FFS.2012", rep("Person_year_FFS.2012", length(healthnames)-1))
healthoutcomes = length(healthnames)

attharray = array(NA, c(healthoutcomes, 5))
dimnames(attharray) = list(healthnames, c("Mean", "SD", "2.5%", "97.5%", "P(<0)"))

edearray = attharray
eaearray = attharray
nedearray = attharray
neaearray = attharray
simslist = list()
nsamp = rep(NA, healthoutcomes)

n = dim(dat)[[1]]

for (hlth in 1:length(healthnames)){
sims = getsims("log", healthoutname = names(healthnames[hlth]), healthout = healthnames[hlth], denomname = denominators[hlth], 
               1.0, -1.0, resultsname = paste(whichresults, "pstratamod_", healthnames[hlth], ".Rda", sep=""), temp = FALSE)

simslist[[hlth]] = sims
nsamp[hlth] = sims[["nsamp"]]

atth = sims[["atth"]][1:(sims$nsamp-nburn)]
ede  = sims[["ede"]][1:(sims$nsamp-nburn)]
nede = sims[["nede"]][1:(sims$nsamp-nburn)]
eae  = sims[["eae"]][1:(sims$nsamp-nburn)]
neae = sims[["neae"]][1:(sims$nsamp-nburn)]

attharray[hlth,"Mean"] = mean(atth)
attharray[hlth,"SD"] = sd(atth)
attharray[hlth,c("2.5%", "97.5%")] = quantile(atth, c(.025, .975))
attharray[hlth,"P(<0)"] = sum(atth<0)/length(atth)

edearray[hlth,"Mean"] = mean(ede, na.rm=TRUE)
edearray[hlth,"SD"] = sd(ede, na.rm=TRUE)
edearray[hlth,c("2.5%", "97.5%")] = quantile(ede, c(.025, .975), na.rm=TRUE)
edearray[hlth,"P(<0)"] = sum(ede<0, na.rm=TRUE)/length(ede)
nedearray[hlth, "Mean"] = mean(nede)
nedearray[hlth, "SD"] = sd(nede)
nedearray[hlth, "P(<0)"] = sum(nede==0)

eaearray[hlth,"Mean"] = mean(eae, na.rm=TRUE)
eaearray[hlth,"SD"] = sd(eae, na.rm=TRUE)
eaearray[hlth,c("2.5%", "97.5%")] = quantile(eae, c(.025, .975), na.rm=TRUE)
eaearray[hlth,"P(<0)"] = sum(eae<0, na.rm=TRUE)/length(eae)
neaearray[hlth, "Mean"] = mean(neae)
neaearray[hlth, "SD"] = sd(neae)
neaearray[hlth, "P(<0)"] = sum(neae==0)

}#hlth

print("Table summarizing average effects on health outcomes:")
print(round(attharray,3))
write.table(round(attharray,3), file = "atth_tab", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=TRUE, quote=FALSE)

print("Table summarizing average dissociative effects:")
print(round(edearray,3))
write.table(round(edearray,3), file = "ede_tab", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=TRUE, quote=FALSE)
print("Table summarizing number of locations used to estimate dissociative effects:")
print(round(nedearray,3))

print("Table summarizing average associative effects:")
print(round(eaearray,3))
write.table(round(eaearray,3), file = "eae_tab", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=TRUE, quote=FALSE)
print("Table summarizing number of locations used to estimate associative effects:")
print(round(neaearray,3))

####### Line Plots of 95% Posterior Intervals #############
## ----  Pick which outcomes to display
displayoutcomes = healthnames
displaynames = fullhealthnames

# pdf("20161031_med.pdf", width = 12)
par(mfrow = c(1,1))
att_x = seq(1, by=4, length = length(displayoutcomes))
ede_x = seq(2, by=4, length = length(displayoutcomes))
eae_x = seq(3, by=4, length = length(displayoutcomes))

xlims = c(min(att_x), max(eae_x))
ylims = c(-6, 6)

plot(xlims, ylims, axes=FALSE, ylab="Causal Effect", xlab="", type = "n")
lines(c(-1000,1000), c(0,0), lty = 2)

points(att_x, attharray[displayoutcomes, "Mean"], pch = 16, col = 1)
arrows(att_x, attharray[displayoutcomes, "2.5%"], att_x, attharray[displayoutcomes, "97.5%"], lwd=2, length=0, col= 1)

points(ede_x, edearray[displayoutcomes, "Mean"], pch = 16, col = 2)
arrows(ede_x, edearray[displayoutcomes, "2.5%"], ede_x, edearray[displayoutcomes, "97.5%"], lwd=2, length=0, col= 2)

points(eae_x, eaearray[displayoutcomes, "Mean"], pch = 16, col = 3)
arrows(eae_x, eaearray[displayoutcomes, "2.5%"], eae_x, eaearray[displayoutcomes, "97.5%"], lwd=2, length=0, col= 3)

axis(side = 2, at = seq(min(ylims), max(ylims), by = 1), labels = round(seq(min(ylims), max(ylims), by = 1), 1))

mtext(displaynames, side = 1, at = ede_x)

legend("topright", c("Overall Average Effect", "Average Dissociative Effect", "Average Associative Effect"), bty = "n", lty = c(1,1,1), pch = c(16,16,16), col = c(1,2,3))
# dev.off()
```

# Sensitivity analysis without pruning
```{r, echo=FALSE}
dat_sens = fulldat
dat_sens$outofrange=0

## Quantiles
cuts = c(0.25, 0.5, 0.75)
dat_sens$pscat[dat_sens$outofrange==0] = as.numeric(cut(dat_sens$ps[dat_sens$outofrange==0], breaks=c(0,cuts,1), include.lowest=TRUE))

print("Table showing number of attainment, nonattainment locations in each propensity score group:")
print(with(dat_sens, table(a,pscat)))

# pdf("balplot_noprune.pdf", height=10, width=10)
bal = mybalcheck(covariates=as.matrix(dat_sens[, printvars, with = FALSE]), treatment=dat_sens$a, propensity=dat_sens$ps, strata=dat_sens$pscat, absolute.es=FALSE, xlim=c(-1,2), outofrange=dat_sens$outofrange, universal.psd=FALSE, plot.strata=FALSE, covnames = covnames)
# dev.off()

stratabal = abs(bal$effect.sizes[, c("a", "b", "c", "d")])

### output a list of variables to include in the models for residual adjustment 
#### include if: 1) average bal > 0.25 or any stratum greater than 0.5
whichresid = matrix(0, nrow = dim(stratabal)[[1]], ncol = dim(stratabal)[[2]])
dimnames(whichresid) = dimnames(stratabal)
residtoobig = ( abs(stratabal) > 0.5)
whichresid[residtoobig] = stratabal[residtoobig]
checkresid = cbind(whichresid, apply(stratabal, 1, mean))
includevar = abs(checkresid[, dim(stratabal)[[2]]+1]) > 0.25 | rowMeans(abs(checkresid[, 1:(dim(stratabal)[[2]])]) ) > 0
cbind(checkresid, includevar)
varnames = dimnames(checkresid)[[1]][includevar]
vars_with_imbalance_noprune = printvars[(covnames %in% varnames) & !(covnames %in% c("Mortality", "All CVD", "Respiratory", "COPD", "CV Stroke", "Heart Failure", "HRD", "Ischemic Heart Disease", "Peripheral Vascular Disease", "Respiratory Tract Infection"))]
save(vars_with_imbalance_noprune, file = "vars_with_imbalance_noprune.RData")
print("Variables with residual imbalance that are included for direct covariate adjustment:")
print(vars_with_imbalance_noprune)

```

## Sensitivity Analysis: Pollution Outcomes
```{r, echo=FALSE}
dat = dat_sens

resultname = paste(whichresults_pollution_noprune, "pollutionmodel.RData", sep = "")
load(resultname)

nsamp = dim(mod$samples)[[1]] 
print(paste("The number of MCMC iterations (after burn in) for the sensitivity analysis of pollution outcomes =", (nsamp - nburn)))
y0out = exp(mod$y0[seq(nburn,nsamp,thin),])
y1out = exp(mod$y1[seq(nburn,nsamp,thin),])

ceymat = y1out-y0out
atty = rowMeans(ceymat[, dat$a==1])

print(paste("The posterior mean effect on ambient PM2.5 = ", round(mean(atty), 3)))
print(paste("and 95% posterior interval = (", round(quantile(atty, c(.025)), 3), ",", round(quantile(atty, c(.975)),3), ")"))
print(paste("The posterior probability that ATTy < 0 = ", round(sum(atty<0)/length(atty),3)))
```

## Sensitivity Analysis: Health Outcomes
```{r, echo=FALSE}
healthnames = c("mort", "copd", "cvstroke", "hf", "hrd", "ihd", "pvd", "rti")    
names(healthnames) = c("total_death_FFS_MA.2012",
                       "COPD.2012", "CV_stroke.2012", "HF.2012", "HRD.2012", "IHD.2012",
                       "PVD.2012", "RTI.2012") 
fullhealthnames = c("Mortality", "COPD", "CV Stroke", "Heart Fail.", "Heart Rhythm Disord.", "IHD", "PVD", "Resp. Tract Infect.")
denominators = c("Tot_den_for_death_MA_FFS.2012", rep("Person_year_FFS.2012", length(healthnames)-1))
healthoutcomes = length(healthnames)

attharray = array(NA, c(healthoutcomes, 5))
dimnames(attharray) = list(healthnames, c("Mean", "SD", "2.5%", "97.5%", "P(<0)"))

edearray = attharray
eaearray = attharray
nedearray = attharray
neaearray = attharray
simslist = list()
nsamp = rep(NA, healthoutcomes)

n = dim(dat)[[1]]

for (hlth in 1:length(healthnames)){
sims = getsims("log", healthoutname = names(healthnames[hlth]), healthout = healthnames[hlth], denomname = denominators[hlth], 
               1.0, -1.0, resultsname = paste(whichresults_noprune, "pstratamod_", healthnames[hlth], ".Rda", sep=""), temp = FALSE)

simslist[[hlth]] = sims
nsamp[hlth] = sims[["nsamp"]]

atth = sims[["atth"]][1:(sims$nsamp-nburn)]
ede  = sims[["ede"]][1:(sims$nsamp-nburn)]
nede = sims[["nede"]][1:(sims$nsamp-nburn)]
eae  = sims[["eae"]][1:(sims$nsamp-nburn)]
neae = sims[["neae"]][1:(sims$nsamp-nburn)]

attharray[hlth,"Mean"] = mean(atth)
attharray[hlth,"SD"] = sd(atth)
attharray[hlth,c("2.5%", "97.5%")] = quantile(atth, c(.025, .975))
attharray[hlth,"P(<0)"] = sum(atth<0)/length(atth)

edearray[hlth,"Mean"] = mean(ede, na.rm=TRUE)
edearray[hlth,"SD"] = sd(ede, na.rm=TRUE)
edearray[hlth,c("2.5%", "97.5%")] = quantile(ede, c(.025, .975), na.rm=TRUE)
edearray[hlth,"P(<0)"] = sum(ede<0, na.rm=TRUE)/length(ede)
nedearray[hlth, "Mean"] = mean(nede)
nedearray[hlth, "SD"] = sd(nede)
nedearray[hlth, "P(<0)"] = sum(nede==0)

eaearray[hlth,"Mean"] = mean(eae, na.rm=TRUE)
eaearray[hlth,"SD"] = sd(eae, na.rm=TRUE)
eaearray[hlth,c("2.5%", "97.5%")] = quantile(eae, c(.025, .975), na.rm=TRUE)
eaearray[hlth,"P(<0)"] = sum(eae<0, na.rm=TRUE)/length(eae)
neaearray[hlth, "Mean"] = mean(neae)
neaearray[hlth, "SD"] = sd(neae)
neaearray[hlth, "P(<0)"] = sum(neae==0)

}#hlth

print(paste("The number of MCMC iterations (after burn in) for the sensitivity analysis of", healthnames, "was", (nsamp -nburn )))

print("Table summarizing average effects on health outcomes:")
print(round(attharray,3))
write.table(round(attharray,3), file = "atth_tab_noprune", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=TRUE, quote=FALSE)

print("Table summarizing average dissociative effects:")
print(round(edearray,3))
write.table(round(edearray,3), file = "ede_tab_noprune", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=TRUE, quote=FALSE)

print("Table summarizing average associative effects:")
print(round(eaearray,3))
write.table(round(eaearray,3), file = "eae_tab_noprune", sep="&", eol="\\\\ \n", row.names=TRUE, col.names=TRUE, quote=FALSE)

####### Line Plots of 95% Posterior Intervals #############
## ----  Pick which outcomes to display
displayoutcomes = healthnames
displaynames = fullhealthnames

# pdf("20161031_med_noprune.pdf", width = 12)
par(mfrow = c(1,1))
att_x = seq(1, by=4, length = length(displayoutcomes))
ede_x = seq(2, by=4, length = length(displayoutcomes))
eae_x = seq(3, by=4, length = length(displayoutcomes))

xlims = c(min(att_x), max(eae_x))
ylims = c(-6, 6)

plot(xlims, ylims, axes=FALSE, ylab="Causal Effect", xlab="", type = "n")
lines(c(-1000,1000), c(0,0), lty = 2)

points(att_x, attharray[displayoutcomes, "Mean"], pch = 16, col = 1)
arrows(att_x, attharray[displayoutcomes, "2.5%"], att_x, attharray[displayoutcomes, "97.5%"], lwd=2, length=0, col= 1)

points(ede_x, edearray[displayoutcomes, "Mean"], pch = 16, col = 2)
arrows(ede_x, edearray[displayoutcomes, "2.5%"], ede_x, edearray[displayoutcomes, "97.5%"], lwd=2, length=0, col= 2)

points(eae_x, eaearray[displayoutcomes, "Mean"], pch = 16, col = 3)
arrows(eae_x, eaearray[displayoutcomes, "2.5%"], eae_x, eaearray[displayoutcomes, "97.5%"], lwd=2, length=0, col= 3)

axis(side = 2, at = seq(min(ylims), max(ylims), by = 1), labels = round(seq(min(ylims), max(ylims), by = 1), 1))

mtext(displaynames, side = 1, at = ede_x)
# dev.off()

```