### 2020-11-18 MSE327/EDA_outcome.Rmd
modified histograms a bit

---
title: "MSE 327 Project"
author: "Davide Giovanardi"
date: "11/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
setwd("/Users/davidegiovanardi/Desktop/MSE327/PM2.5-Nonattainment/MSE327/")
load("pmanalysis.RData")
```

# EDA
## Outcomes
### Ambient PM2.5 2010-2012

```{r EDA}
# Check that it matches Table 1
cat(mean(na.omit(dat[dat$a == 0, "pmfu"])))
cat(mean(na.omit(dat[dat$a == 1, "pmfu"])))

cat(mean(na.omit(dat[dat$a == 0, "HF.2012"])))

clean_df <- function(dat, outcome) {
  df <- na.omit(dat[,c("a",outcome)])
  return(data.frame(df))
}

df <- clean_df(dat, "pmfu")

df$Area <- c("attainment", "non-attainment")[(df$a)+1]

plot1 <- ggplot(df, aes(pmfu, color = Area, fill = Area)) +
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5, bins = 25) +
  geom_density(alpha = 0.25) +
  theme_minimal() + 
  ggtitle("Outcome: Ambient PM2.5 2010-2012") +
  xlab("Ambient PM2.5 2010/12")
```

```{r}
# Check that it matches Table 1
cat(mean(na.omit(dat[dat$a == 0, "pmbase2002_2004"])))
cat(mean(na.omit(dat[dat$a == 1, "pmbase2002_2004"])))

cat(mean(na.omit(dat[dat$a == 0, "HF.2012"])))

clean_df <- function(dat, outcome) {
  df <- na.omit(dat[,c("a",outcome)])
  return(data.frame(df))
}

df_1 <- clean_df(dat, "pmbase2002_2004")

df_1$Area <- c("attainment", "non-attainment")[(df_1$a)+1]

plot2 <- ggplot(df_1, aes(pmbase2002_2004, color = Area, fill = Area)) +
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5, bins = 25) +
  geom_density(alpha = 0.25) +
  theme_minimal() + 
  ggtitle("Outcome: Ambient PM2.5 2002-2004") + 
  xlab("Ambient PM2.5 2002/04")
```

```{r}
# Check that it matches Table 1

clean_df_1 <- function(dat, outcome1, outcome2) {
  df <- na.omit(dat[,c("a",outcome1, outcome2)])
  return(data.frame(df))
}

df_2 <- clean_df_1(dat, "pmfu", "pmbase2002_2004")

df_2$Area <- c("attainment", "non-attainment")[(df_2$a)+1]
df_2$dif <- df_2$pmbase2002_2004 - df_2$pmfu

plot3 <- ggplot(df_2, aes(dif, color = Area, fill = Area)) +
  geom_histogram(aes(y=..density..), position="identity", alpha=0.5, bins = 25) +
  geom_density(alpha = 0.25) +
  theme_minimal() + 
  ggtitle("Outcome: Change in Ambient PM2.5 between 2002/04 and 2010/12") + 
  xlab("Ambient [PM2.5] 2002/04 - Ambient [PM2.5] 2010/12")
```

```{r}
par(mfrow = c(2, 2))
plot1
plot2
plot3
```